### 2022-04-23 write-up/protocol.Rmd
add draft protocol; add styles; add bibliography

---
title: "Comparative effectiveness of BNT162b2 versus Moderna booster doses in England: protocol"
output:
  html_document:
    self_contained: yes
  pdf_document:
    toc: no
  bookdown::html_document2:
    number_sections: no
    toc: no
  rtf_document:
    toc: no
  word_document: default
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r setup, include=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')

# remotes::install_github("https://github.com/wjchulme/osutils")
# library('osutils')

# where are the outputs (ie the inputs for this manuscript!) saved?
output_dir_os <- here("output", "release-objects")
#output_dir_os <- here("released_output", "release-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("write-up", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

## Import custom user functions from lib
source(here("lib", "functions", "utility.R"))

## Import design elements
source(here("lib", "design", "design.R"))

```

\newpage

## Objectives

To assess the compare the two most widely used COVID-19 booster vaccinations in England, Pfizer-BioNTech (BNT162b2) and Moderna (mRNA-1273) in their effectiveness against COVID-19-related outcomes.

## Overview

The UK COVID-19 vaccination programme delivered its first "booster" doses in September 2021, initially in groups at high risk of severe disease then across the adult population. The BNT162b2 Pfizer-BioNTech vaccine was used initially, with Moderna mRNA-1273 subsequently also used. The concurrent administration BNT162b2 and Moderna vaccines for boosting in England enables a direct comparison of their effectiveness. We used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to compare the effectiveness of BNT162b2 versus Moderna in boosted adults between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`, during which time both vaccine brands were being used.

## Methods

## Data source

This study will use the [OpenSAFELY platform](https://opensafely.org/), providing secure access to the OpenSAFELY-TPP database containing primary care records managed by the GP software provider TPP, roughly 40% of English primary care practices in England. These recorded are linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS).

## Study population

We will consider all adults aged $\ge$ 18 years who received a booster dose of BNT162b2 or Moderna between `r format(as.Date(study_dates$studystart_date), "%e %B")` and `r format(as.Date(study_dates$studyrend_date), "%e %B %Y")` inclusive (the "recruitment period"). People will be eligible for matching if:

* they are registered at a GP practice using TPP's SystmOne clinical information system at the time of boosting;
* they received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers);
* they are not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care;
* there is no evidence of SARS-CoV-2 infection or COVID-19 disease within the 90 days prior to booster dose;
* they are not undergoing an unplanned hospital admission at the time of boosting;
* they have complete information on sex, ethnicity, deprivation, and NHS region.

## Exposure

The exposure of interest is the brand of COVID-19 vaccine administered for the third or booster dose; either BNT162b2 or Moderna. Other vaccine brands were used much less frequently &mdash; for example if there were contraindications or previous reactions to BNT162b2 or Moderna brands &mdash; and so will not be considered in this study.

## Outcomes

The following primary outcomes were considered: 
* Positive SARS-CoV-2 test, identified using SGSS testing records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection. 
* COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc. This outcome also includes COVID-19 admissions (see below) to pick up more severe disease episodes that may have bypassed A&E. 
* COVID-19 hospital admission, identified using HES in-patient hospital records with U07.1 or U07.2 ICD-10 reason for admission codes. 
* COVID-19 death. Deaths were classified as from COVID-19 if deaths with the U07.1 or U07.2 ICD-10 codes were mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
* Non-COVID-19 death.

Additional secondary outcomes:
* COVID-19 A&E attendances that resulted in a hospital admission, according to the discharge location. This can be used as a surrogate for hospital admission 
* COVID-19 ICU / critical care admission

ICU admission and mortality are expected to be low in this group of healthy vaccine recipients, and there may be insufficient statistical power to assess these outcomes reliably.

## Study duration

Time-zero is the day of booster dose receipt. Booster recipients will be followed from the this day until the first occurrence of:

* The outcome of interest
* death
* practice de-registration
* 10 weeks since boosting
* fourth vaccine dose
* `r format(as.Date(study_dates$followupend_date), "%e %B %Y")`.

Outcome events occurring on the day of boosting will be considered to have occurred _after_ boosting, one day after time-zero (i.e., boosting is assumed to have occurred at the start of the day, and all other events are assumed to have occurred at the end of the day). This is necessarily true for death and likely true for hospitalisation outcomes. For positive SARS-CoV-2 tests the order of events is less clear, but given the policy of a recent positive test precluding vaccination, this is a reasonable assumption.

## Matching

BNT162b2 booster recipients will be matched with Moderna booster recipients on one or more of the following characteristics, calculated as at the day of boosting:

Essential matching criteria:
* date of booster dose
* primary course vaccine brand
* date of second vaccine dose, within 7 days
* evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission)
* clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither)
* age, within 3 years, and age groups used by JCVI for prioritisation
* Geographical region, one of the below in decreasing order of precision:
  * Lower Layer Super Output Area (LSOA)
  * Middle Layer Super Output Area (MSOA)
  * NHS Strategic Health Partnership (STP)
  * NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West)

Additional matching criteria if possible:
* English Index of Multiple Deprivation (IMD), within X days (tbd)
* ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census);
* existing morbidity count: diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease
* learning disabilities
* severe mental illness
* the number of SARS-CoV-2 tests in the 6 months prior to boosting
* elective in-hospital episode at the time of boosting

The final specification of matching characteristics will depend on the time required to complete the matching run (matching complexity will be reduced if it is infeasibly long), and number of candidates successfully matched (matching complexity will be reduced if success rate is too low).

We will use one-to-one matching without replacement initially, but may revise to include many-to-many matching to increase power if required. However, many-to-many matching is only straight-forward with pure exact matching. If any variables are matched using calipers, then nearest-neighbour or optimal matching will be required instead.

Deprivation (IMD) is defined within LSOAs (lower-layer super output areas). Each MSOA (middle-layer super output area) is composed of a small number of LSOAs. MSOA is the smallest geographical unit available within OpenSAFELY-TPP. If MSOA is used as a matching variable then deprivation will add little additional information, and may be poorly balanced, and so should not be used. If MSOA is not used, IMD should be used if possible.

## Statistical Analysis

We estimated Kaplan-Meier (KM) cumulative incidence curves separately by vaccine brand, and compared brands using the 10-week risk differences and risk ratios.

Confidence limits for the KM estimate will be derived from complimentary log-log transformation.
Confidence limits for the risk difference will be derived from the square root of the sum of squares the treatment-specific KM standard errors. The KM standard errors will use Greenwood's formula.
Confidence limits for the relative risk??

Alternatively, we could boostrap.

<!-- [The primary estimand is the average treatment effect in the recruited population. To estimate this, expected cumulative incidence rates are estimated for each participant assuming treatment was received by everyone in both treatment groups at time zero. These curves are then averaged over all participants to give the marginal cumulative incidence. This is repeated assuming nobody is treated in either group, and their difference taken. Calculating standard errors for these quantities is challenging, whether approached analytically (due to complex variance components) or computationally (via bootstrapping, due to the large sample sizes involved) and are therefore not reported. ] -->

## Subgroup analyses

We also estimated comparative effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.

## Software, code, and reproducibility

Data management and analyses will be conducted in Python version 3.8.10 and R version 4.0.2. All code will be shared openly for review and re-use under MIT open license at <https://github.com/opensafely/comparative-booster>.

This study will follow the STROBE-RECORD reporting guidelines.

## Disclosure control

Any reported figures based on counts below 6 must be redacted or rounded for disclosure control. To reduce secondary disclosure risk (e.g., by using two unredacted values to derive a redacted value), rounding is preferred.

Kaplan-Meier survival estimates are rounded upwards by deferring event-times until there are at least 7 events, so that each KM "step" is based on at least 7 events. The same principle is applied to KM confidence limits.

# Results

## Study population and matching

`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0") %>% pull(n) %>% sum)` adults registered at a TPP practice received a booster vaccination of either BNT162b2 (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="pfizer") %>% pull(n))`) or Moderna (`r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c0", vax3_type=="moderna") %>% pull(n))`) during the study period, with `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(pct_all))`) and `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c6", vax3_type="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(match_flowchart, crit=="c6", vax3_type="moderna") %>% pull(pct_all))`) eligible for matching. `r label_number(1, big.mark=",")(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(n))` were successfully matched in each group, `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="pfizer") %>% pull(pct_step))` and `r label_percent(0.1)(filter(match_flowchart, crit=="c7", vax3_type="moderna") %>% pull(pct_step))` respectively (Table 1, Figure 1).

#### Table 1: inclusion criteria

```{r table-flowchart, echo=FALSE, warning=FALSE, message=FALSE}

match_flowchart %>%
  select(-crit) %>%
   pivot_wider(
    id_cols = c(criteria, vax3_type),
    names_from = vax3_type,
    values_from = c(n, n_exclude, pct_all)
  ) %>%
  gt() %>%
  cols_label(
    criteria = "",

    n_pfizer = "N",
    n_moderna = "N",

    n_exclude_pfizer = "N excluded",
    n_exclude_moderna = "N excluded",

    pct_all_pfizer = "% remaining",
    pct_all_moderna = "% remaining",
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("pfizer")
  ) %>%
   tab_spanner(
    label = "Moderna mRNA-1273",
    columns = ends_with("moderna")
  ) %>%
  # tab_spanner(
  #   label = "mRNA-1273",
  #   columns = ends_with("moderna")
  # ) %>%
  fmt_percent(
    columns = starts_with(c("pct_")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("n_")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "criteria")
  )
```

\newpage 
#### Figure 1: Cumulative matching success

```{r, figure-matchingsuccess, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))

xmin <- min(match_coverage$vax3_date )
xmax <- max(match_coverage$vax3_date )+1

plot_coverage_cumuln <-
  match_coverage %>%
  mutate(
    treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
    status_descr = fct_recoderelevel(as.character(status), recoder$status),
    cumuln = cumuln*((treatment*2) - 1)
  ) %>%
  ggplot()+
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=paste0(treatment,status_descr),
      fill=treatment_descr,
      alpha=status_descr,
      colour=NULL
    ),
    position=position_stack(reverse=TRUE),
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(match_coverage$vax3_date, "1 month")),
    limits = c(xmin-1, NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = ~scales::label_number(accuracy = 1, big.mark=",")(abs(.)),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2")+
  scale_colour_brewer(type="qual", palette="Set2")+
  scale_alpha_discrete(range= c(0.8,0.4))+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "bottom"
  )+
  NULL

plot_coverage_cumuln
cat("  \n\n")
```

\newpage

#### Table 2: baseline characteristics

```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
table1 <- 
  match_table1 %>%
  #filter(outcome %in% outcomes) %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(match_table1by$by_col, match_table1by$by_chr))

table1 %>%
  gt(groupname_col="treatment_descr") %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
     label=md("Characteristic")
  ) %>%
  cols_hide(c("variable", "var_label", "order"))
```

## Estimated comparative effectiveness

### Main analysis

#### Table 3: Kaplan-Meier cumulative incidence estimates

```{r table-km, echo=FALSE, warning=FALSE, message=FALSE}

km_contrasts_overall %>%
  transmute(
    outcome_descr,
    subgroup_descr,
    subgroup_level_descr,
    n.atrisk=n.atrisk_0,
    cml.atrisk_0=label_number(1)(cml.atrisk_0/7), 
    cml.atrisk_1=label_number(1)(cml.atrisk_1/7),
    n.event_0, 
    n.event_1,
    riskCI_0 = paste0(label_number(0.01)(risk_0), " (", label_number(0.01)(risk.ll_0), "-", label_number(0.01)(risk.ul_0), ")"),
    riskCI_1 = paste0(label_number(0.01)(risk_1), " (", label_number(0.01)(risk.ll_1), "-", label_number(0.01)(risk.ul_1), ")"),
    kmrdCI = paste0(label_number(0.01)(kmrd), " (", label_number(0.01)(kmrd.ll), "-", label_number(0.01)(kmrd.ul), ")"),
    kmrr = paste0(label_number(0.01)(kmrr)),
  ) %>%
  gt(groupname_col="outcome_descr") %>%
  cols_label(

    outcome_descr = "Outcome",
    subgroup_descr = "Subgroup",
    subgroup_level_descr = "Subgroup",

    n.atrisk = "N per group",

    cml.atrisk_0 = "Person-weeks", 
    cml.atrisk_1 = "Person-weeks",

    n.event_0 = "Events",
    n.event_1 = "Events",

    riskCI_0 = "KM risk (95% CI)",
    riskCI_1 = "KM risk (95% CI)",

    kmrdCI = "Risk-difference (95% CI)",
    kmrr = "Relative-risk"
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("_0")
  ) %>%
   tab_spanner(
    label = "Moderna",
    columns = ends_with("_1")
  )  %>%
   tab_spanner(
    label = "Comparison",
    columns = starts_with("km")
  )

```

#### Figure 3: Kaplan-Meier cumulative incidence estimates

```{r, figure-km, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
plot_km <- km_estimates %>%
  filter(
    subgroup=="all",
    outcome %in% outcomes
  ) %>%
  ggplot(aes(group=treatment_descr, colour=treatment_descr, fill=treatment_descr)) +
  geom_step(aes(x=time, y=1-surv))+
  geom_rect(aes(xmin=time, xmax=leadtime, ymin=1-surv.ll, ymax=1-surv.ul), alpha=0.1, colour="transparent")+
  facet_grid(rows=vars(outcome_descr))+
  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
  scale_x_continuous(breaks = seq(0,600,14))+
  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
  coord_cartesian(xlim=c(0, NA))+
  labs(
    x="Days",
    y="Cumulative incidence",
    colour=NULL,
    title=NULL
  )+
  theme_minimal()+
  theme(
    axis.line.x = element_line(colour = "black"),
    panel.grid.minor.x = element_blank(),
    legend.position=c(.05,.95),
    legend.justification = c(0,1),
  )

plot_km
```

# Supplement

## JCVI vaccine priority groups

The Joint Committee on Vaccine and Immunisation (JCVI) priority groups were defined as follows:

```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE}

tibble(
  `Priority group` = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10a", "10b"),
  `Description` = c(
    "Residents in a care home for older adults\nStaff working in care homes for older adults",
    "All those 80 years of age and over\nFrontline health and social care workers",
    "All those 75-79 years of age",
    "All those 70-74 years of age\nIndividuals aged 16-69 in a high risk group",
    "All those 65-69 years of age",
    "Adults aged 16-64 years in an at-risk group",
    "All those 60-64 years of age",
    "All those 55-59 years of age",
    "All those 50-54 years of age",
    "All those 40-49 years of age",
    "All those 18-39 years of age"
  )
) %>%
mutate(Description = str_replace_all(Description, "\n", "<br>")) %>% 
gt() %>% 
fmt_markdown(columns = TRUE)

```

See original priority groups here: https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1007737/Greenbook_chapter_14a_30July2021.pdf#page=15
See revised priority groups here: https://www.england.nhs.uk/coronavirus/wp-content/uploads/sites/52/2021/07/C1327-covid-19-vaccination-autumn-winter-phase-3-planning.pdf

Note that we excluded care home residents and health care workers in our analysis, so members of JCVI group 1 are not included and JCVI group 2 includes only those aged 80 and over. The original priority group list has 9 groups, with a 10th group implicitly defined as "everybody else". Here we explicitly define this group, and split into two (10a and 10b) because of the earlier booster eligibility in the 40-49 group from 15 November 2021 onwards. (https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds)

\newpage

## References