### 2022-07-06 write-up/supplement.Rmd

---
title: "Supplementary materials for Comparative effectiveness of BNT162b2 versus mRNA-1273 booster doses in England"
output:
  html_document:
    self_contained: yes
  pdf_document:
    toc: no
  bookdown::html_document2:
    number_sections: no
    toc: no
  rtf_document:
    toc: no
  word_document: default
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r setup, include=FALSE, message=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')

# remotes::install_github("https://github.com/wjchulme/osutils")
# library('osutils')

# where are the outputs (ie the inputs for this manuscript!) saved?
#output_dir_os <- here("output", "release-objects")
output_dir_os <- here("released-output", "release-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("write-up", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

## Import custom user functions from lib
source(here("lib", "functions", "utility.R"))

## Import design elements
source(here("lib", "design", "design.R"))

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidadmittedproxy1",
      #"covidcritcare",
      "coviddeath",
      #"noncoviddeath"
    ) %>% 
    set_names(., .)

```

```{r match-A, echo=FALSE, warning=FALSE, message=FALSE}

# 
# prematch_table1 <- read_csv(fs::path(output_dir_os, "prematch", "match_table1.csv"))
# prematch_table1by <- read_csv(fs::path(output_dir_os, "prematch", "match_table1by.csv"))

## matching A

matchA_flowchart <-  read_csv(fs::path(output_dir_os, "A", "match_flowchart.csv"))
matchA_coverage <- read_csv(fs::path(output_dir_os, "A", "match_coverage.csv")) %>%
   mutate(
    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
  )

matchA_table1 <- read_csv(fs::path(output_dir_os, "A", "match_table1.csv"))
matchA_table1by <- read_csv(fs::path(output_dir_os, "A", "match_table1by.csv"))
matchA_smd <-  read_csv(fs::path(output_dir_os, "A", "match_smd.csv"))

ciA_estimates <- read_csv(fs::path(output_dir_os, "A", "ci_estimates.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsA_daily <- read_csv(fs::path(output_dir_os, "A", "ci_contrasts_daily.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsA_cuts <- read_csv(fs::path(output_dir_os, "A", "ci_contrasts_cuts.csv")) %>%
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsA_overall <- read_csv(fs::path(output_dir_os, "A", "ci_contrasts_overall.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, ""),

  )

## matching B

matchB_flowchart <-  read_csv(fs::path(output_dir_os, "B", "match_flowchart.csv"))
matchB_coverage <- read_csv(fs::path(output_dir_os, "B", "match_coverage.csv")) %>%
   mutate(
    treatment_descr = fct_recoderelevel(as.character(treatment),  recoder$treatment),
  )

matchB_table1 <- read_csv(fs::path(output_dir_os, "B", "match_table1.csv"))
matchB_table1by <- read_csv(fs::path(output_dir_os, "B", "match_table1by.csv"))
matchB_smd <-  read_csv(fs::path(output_dir_os, "B", "match_smd.csv"))

ciB_estimates <- read_csv(fs::path(output_dir_os, "B", "ci_estimates.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsB_daily <- read_csv(fs::path(output_dir_os, "B", "ci_contrasts_daily.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsB_cuts <- read_csv(fs::path(output_dir_os, "B", "ci_contrasts_cuts.csv")) %>%
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, "")
  )
contrastsB_overall <- read_csv(fs::path(output_dir_os, "B", "ci_contrasts_overall.csv")) %>% 
  mutate(
    outcome_descr = fct_recoderelevel(as.character(outcome),  recoder$outcome),
    subgroup_descr = fct_recoderelevel(subgroup,  recoder$subgroup),
    subgroup_level_descr = replace_na(subgroup_level_descr, ""),

  )

perN <- 1000
perN_format <- label_number(1, 1, big.mark=",")(perN)

label_number_n <- label_number(1, 1, big.mark=",")
label_number_risk <- label_number(0.01, perN, style_negative="minus", big.mark=",")
label_number_rr <- label_number(0.01, 1, style_negative="minus")

```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Table S1: baseline characteristics of booster vaccine recipients eligible for matching

```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
# table1 <- 
#   prematch_table1 %>%
#   group_by(variable) %>%
#   mutate(
#     label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
#   ) %>%
#   mutate(
#     across(
#       starts_with("stat_"),
#       ~if_else(is.na(.x) & row_number()==1, "", .x)
#     )
#   ) %>%
#   ungroup() %>%
#   mutate(
#     order = cumsum(lag(variable, 1, first(variable))!=variable)
#   ) %>%
#   select(
#     variable, var_label, label, starts_with("stat_"), order
#   ) %>%
#   rename(!!!set_names(match_table1by$by_col, match_table1by$by_chr))
# 
# 
# table1 %>%
#   gt(groupname_col="treatment_descr") %>%
#   fmt_markdown(columns = "label") %>%
#   tab_style(
#     style = list(
#       cell_fill(color = "gray96")
#     ),
#     locations = cells_body(
#       rows = order%%2 == 0
#     )
#   ) %>%
#   tab_style(
#     style = list(
#       cell_text(weight = "bold")
#     ),
#     locations = cells_column_labels(everything())
#   ) %>%
#   cols_label(
#      label=md("Characteristic")
#   ) %>%
#   cols_hide(c("variable", "var_label", "order")) %>%
#   tab_options(
#     table.font.size = 8
#   )
# 
# cat("  \n\n")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S1: Standardised Mean Differences between BNT12b2 and mRNA-1273 groups

<!-- For variables that were matched exactly the SMD is zero, and so are not shown. -->

```{r figure-smd, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

match_smd <-
  bind_rows(
    matchA_smd %>% mutate(strategy="A"), 
    matchB_smd %>% mutate(strategy="B"), 
  )

plot_smd <-
  match_smd %>%
  filter(
    !(var_label %in% c("Primary vaccine course", "Region", "Prior documented SARS-CoV-2 infection", "JCVI clinical risk group", "IMD"))
  ) %>%
  mutate(
    level=fct_rev(fct_inorder(level))
  ) %>%
  droplevels() %>%
  group_by(strategy) %>%
  mutate(
    variable = fct_inorder(variable),
    variable_card = as.numeric(variable)%%2,
    variable_levels = replace_na(as.character(variable_levels), ""),
  ) %>%
  arrange(strategy, variable) %>%
  mutate(
    level = fct_rev(fct_inorder(str_replace(paste(variable, variable_levels, sep=": "),  "\\:\\s$", ""))),
    cardn = row_number()
  ) %>%
  ggplot()+
  geom_point(aes(x=smd, y=level, group=strategy, colour=strategy), position=position_dodge(width=0.3))+
  geom_linerange(aes(xmax=smd, xmin=0, y=level, group=strategy, colour=strategy), position=position_dodge(width=0.3))+
  geom_rect(aes(alpha = variable_card, ymin = rev(cardn)-0.5, ymax =rev(cardn+0.5), group=strategy), xmin = -Inf, xmax = Inf, fill='grey', colour="transparent") +
  scale_alpha_continuous(range=c(0,0.1), guide=FALSE)+
  scale_colour_brewer(type="qual", palette="Set2")+
  labs(
    x="Standardised mean difference\n <- proportion higher in BNT162b2 | proportion higher in mRNA-1273 -> ",
    y=NULL,
    alpha=NULL,
    colour= "Matching strategy"
  )+
  theme_minimal(base_size=10) +
  theme(
    legend.position = "bottom",
    strip.placement = "outside",
    strip.background = element_rect(fill="transparent", colour="transparent"),
    strip.text.y.left = element_text(angle = 0, hjust=1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.spacing = unit(0, "lines")
  )

plot_smd

```

\newpage

## Comparative effectiveness estimates, matching strategy A

### Cumulative incidence

```{r, function-ci, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}

ci_theme <- function(...) {
  theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
      ...
    )
}

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_ci_subgroup <- function(subgroup){

  subgroup0 <- subgroup

  if(subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_ci <- ciB_estimates %>%
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    mutate(
      colgroup = interaction(treatment_descr, subgroup_level_descr, sep=", "),
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10)
    ) %>%
    #group_by(colgroup, outcome_descr) %>%
    ggplot(aes(group=colgroup, colour=treatment_descr, fill=treatment_descr, alpha=treatment_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    geom_step(aes(x=lagtime, y=risk, linetype=treatment_descr))+
    #geom_segment(aes(x=lagtime, xend=time, y=risk, yend=risk))+
    #geom_segment(aes(x=lagtime, xend=lagtime, y=lag(risk,1,0), yend=risk))+
    geom_rect(aes(xmin=lagtime, xmax=time, ymin=risk.ll, ymax=risk.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    #scale_color_manual(values=colourvalues, na.value="grey") +
    #scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number_risk)+
    scale_alpha_discrete(range=c(0.3,0.9))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Days",
      y=NULL,
      #title=glue("Cumulative incidence per {perN_format}"),
      colour=NULL,
      alpha=NULL,
      linetype=NULL
    )+
    guides(linetype="none", fill="none", alpha="none")+
    theme_minimal(base_size = 9)+
      theme(
        # legend.position = c(0.05,0.1),
        # legend.justification = c(0,0),
        legend.position = "bottom",
        #legend.direction = "vertical",
        axis.text.x.top=element_text(hjust=0),

        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),
        strip.placement="outside",
        strip.text.y.left = element_text(angle=0),

        panel.border = element_blank(),
        panel.spacing = unit(0.8, "lines"),

      )+
      NULL

  plot_coxhr <- contrastsB_cuts %>%
    # equivalent to the hazard
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    group_by(outcome_descr, subgroup_level_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    geom_segment(aes(x=period_start, xend=period_end, y=coxhr, yend=coxhr))+
    geom_segment(aes(x=period_start, xend=period_start, y=lag(coxhr), yend=coxhr))+
    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=coxhr.ll, ymax=coxhr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y")+
    # scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    # scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_color_manual(values=colourvalues, na.value="grey") +
    scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
    coord_cartesian(xlim=c(0, NA))+
    colour_guide+
    labs(
      x="Days",
      y="higher risk\n<- BNT162b2 | mRNA 1273 ->",
      title="Hazard ratio",
      colour=NULL
    )+
    theme_minimal(base_size = 9)+
    ci_theme()+
    NULL

  n_subgroups <-
    ciB_estimates %>%
    filter(
      subgroup==subgroup0
    ) %>% 
    `[[`("subgroup_level_descr") %>%
    n_distinct()

  plot_combined <- patchwork::wrap_plots(
    plot_ci, 
    #plot_cird, 
    #plot_cirr, 
    #plot_kmirr,
    plot_coxhr,
    ncol=2,
    widths=c(n_subgroups,1)
  )

  ggsave(
    filename=glue("ciB_{subgroup}.png"),
    path=output_dir_rmd,
    plot=plot_ci,
  )

  plot_ci
}

```

#### Figure S2a: Cumulative incidence estimates per `r perN_format` people

```{r, figure-ci-A-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("all")
```

#### Figure S2b: Cumulative incidence estimates per `r perN_format` people by primary course type

```{r, figure-ci-A-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("vax12_type")
```

#### Figure S2c: Cumulative incidence estimates per `r perN_format` people by age group

```{r, figure-ci-A-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("age65plus")
```

#### Figure S2d: Cumulative incidence estimates per `r perN_format` people by prior infection status

```{r, figure-ci-A-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("prior_covid_infection")
```

#### Figure S2e: Cumulative incidence estimates per `r perN_format` people by clinical vulnerability

```{r, figure-ci-A-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("cev_cv")
```

```{r, function-plot-contrasts, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_contrast_subgroup <- function(subgroup, matchset){

  subgroup0 <- subgroup

  if(subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_data <- get(glue("contrasts{matchset}_daily")) %>%
    filter(
      subgroup==subgroup0,
      outcome %in% outcomes
    ) %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    mutate(
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
      lagcird=lag(cird,1,0),
      lagcirr=lag(cirr,1,1)
    )

  plot_cird <- 
    plot_data %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    geom_segment(aes(x=period_start, xend=period_end, y=cird, yend=cird))+
    geom_segment(aes(x=period_start, xend=period_start, y=lagcird, yend=cird))+
    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=cird.ll, ymax=cird.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_continuous(expand = expansion(mult=c(0,0.01)), label=label_number(0.1, 10000))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Days",
      y=NULL,
      title="Risk difference",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      # legend.position = c(0.05,0.1),
      # legend.justification = c(0,0),
      legend.position = "bottom",
      #legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement="outside",
      strip.text.y.left = element_text(angle=0),

      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
    )+
    NULL

  plot_cirr <- 
    plot_data %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    geom_segment(aes(x=period_start, xend=period_end, y=cirr, yend=cirr))+
    geom_segment(aes(x=period_start, xend=period_start, y=lagcirr, yend=cirr))+
    geom_rect(aes(xmin=period_start, xmax=period_end, ymin=cirr.ll, ymax=cirr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,14), expand = expansion(mult=c(0,0.01)))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Days",
      y=NULL,
      title="Risk ratio",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines")
    )+
    NULL

  plot_combined <- patchwork::wrap_plots(
    plot_cird, 
    plot_cirr, 
    ncol=2
  )

  ggsave(
    filename=glue("contrasts{matchset}_{subgroup}.png"),
    path=output_dir_rmd,
    plot=plot_combined,
  )

  plot_combined
}

```

### Cumulative contrasts

#### Figure S3a: Cumulative risk difference and risk ratio for matching strategy A

```{r, figure-contrasts-A-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("all","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S3b: Cumulative risk difference and risk ratio for matching strategy A by primary course type

```{r, figure-contrasts-A-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("vax12_type","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S3c: Cumulative risk difference and risk ratio for matching strategy A by age group

```{r, figure-contrasts-A-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("age65plus","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S2d: Cumulative risk difference and risk ratio for matching strategy A by prior infection status

```{r, figure-contrasts-A-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("prior_covid_infection","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S3e: Cumulative risk difference and risk ratio for matching strategy A by clinical vulnerability

```{r, figure-contrasts-A-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("cev_cv","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

## Comparative effectiveness estimates, matching strategy B

### Cumulative incidence

#### Figure S4a: Cumulative incidence estimates per `r perN_format` people for matching strategy B

```{r, figure-ci-B-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("all")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S4b: Cumulative incidence estimates per `r perN_format` people for matching strategy B by primary course type

```{r, figure-ci-B-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("vax12_type")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S4c: Cumulative incidence estimates per `r perN_format` people for matching strategy B by age group

```{r, figure-ci-B-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("age65plus")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S4d: Cumulative incidence estimates per `r perN_format` people for matching strategy B by prior infection status

```{r, figure-ci-B-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("prior_covid_infection")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S4e: Cumulative incidence estimates per `r perN_format` people for matching strategy B by clinical vulnerability

```{r, figure-ci-B-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_ci_subgroup("cev_cv")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

### Cumulative contrasts

#### Figure S5a: Cumulative risk difference and risk ratio for matching strategy B

```{r, figure-contrasts-B-all, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("all","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S5b: Cumulative risk difference and risk ratio for matching strategy B by primary course type

```{r, figure-contrasts-B-vax12_type, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("vax12_type","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S5c: Cumulative risk difference and risk ratio for matching strategy B by age group

```{r, figure-contrasts-B-age65plus, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("age65plus","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S5d: Cumulative risk difference and risk ratio for matching strategy B by prior infection status

```{r, figure-contrasts-B-prior_covid_infection, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("prior_covid_infection","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Figure S5e: Cumulative risk difference and risk ratio for matching strategy B by clinical vulnerability

```{r, figure-contrasts-B-cev_cv, echo=FALSE, message=FALSE, warning=FALSE, out.width='100%', out.height='50%', results='asis', out.extra=''}
plot_contrast_subgroup("cev_cv","A")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage