### 2024-01-31 write-up/manuscript/supplement.qmd

---
title: Supplementary materials for Comparative safety and effectiveness of Pfizer BA.4-5 versus Sanofi during the Spring 2023 COVID-19 booster vaccination programme in England
format:
  # html:
  #   toc: false
  #   embed-resources: true
  docx
output-file: supplementary_draft.docx
keep-md: false
execute:
  echo: false
  output: true
  warning: false
  #message: true
knitr:
  opts_chunk: 
    #fig.path: ./
    fig.pos: H
    R.options:
      knitr.graphics.auto_pdf: true
bibliography: references.bib
csl: vancouver.csl
link-citations: true
zotero: true
# header-includes:
# - \usepackage{float} 
# - \floatplacement{figure}{H}
# - \usepackage{caption}
# - \captionsetup[figure]{labelformat=empty}
# - \captionsetup[table]{labelformat=empty}
---

```{r}
#| label: setup

source(here::here("write-up", "manuscript", "setup.R"))
```

```{r}
#| label: process

source(here("write-up", "manuscript", "import-process.R"))
```

```{r}
#| label: figure-function

show_coverage <- function(.cohort){

  xmin <- min(match_coverage$boost_date )
  xmax <- max(match_coverage$boost_date )+1

  data_annotate <- tibble(
      boost_date = date(c("2023-05-01", "2023-05-01", "2023-05-20", "2023-06-07")),
      cumuln = c(1000000, -100000, 0, -10000),
      dailyn = c(50000, -25000, 0, -12000),
      label = c("Sanofi", "Pfizer BA.4-5", "Matched", "Unmatched"),
      colour = c(brewer_pal(type="qual", palette="Set2")(2)[c(2,1)], "grey50", "grey70"),
      vjust = c(-1, 1, 0, 0.5)
    )

  plot_coverage_cumuln <-
    rst[[.cohort]]$match_coverage %>%
    mutate(
     # treatment_descr = fct_recoderelevel(as.character(treatment), recoder$treatment),
      #status_descr = fct_rev(fct_recoderelevel(as.character(status), recoder$status)),
      #status_descr = fct_rev(status_descr),
      cumuln = cumuln*((treatment*2) - 1),
      dailyn = n*((treatment*2) - 1)
    ) %>%
    ggplot()+
    geom_col(
      aes(
        x=boost_date+0.5,
        y=dailyn,
        group=paste0(treatment,status_descr),
        fill=treatment_descr,
        alpha=status_descr,
        colour=NULL
      ),
      position=position_stack(reverse=TRUE),
      width=1
    )+
    geom_rect(xmin=xmin, xmax= xmax+1, ymin=-6, ymax=6, fill="grey", colour="transparent")+
    geom_text(data=data_annotate, aes(x=boost_date, y=dailyn, label=label), colour=data_annotate$colour, vjust=data_annotate$vjust, hjust=1.05, fontface="bold")+
    scale_x_date(
      breaks = seq(xmin, xmax+1, by=14),
      #breaks = unique(lubridate::ceiling_date(match_coverage$boost_date, "1 month")),
      limits = c(xmin-1, NA),
      labels = scales::label_date("%d/%m"),
      expand = expansion(add=1),
      sec.axis = sec_axis(
        trans = ~as.Date(.),
        breaks=as.Date(seq(floor_date(xmin, "month"), ceiling_date(xmax, "month"),by="month")),
        labels = scales::label_date("%B %y")
      )
    )+
    scale_y_continuous(
      labels = ~scales::label_number(accuracy = 1, big.mark=",")(abs(.)),
      expand = expansion(c(0, NA))
    )+
    scale_fill_brewer(type="qual", palette="Set2")+
    scale_colour_brewer(type="qual", palette="Set2")+
    scale_alpha_discrete(range= c(0.8,0.4))+
    labs(
      x="Date",
      y="Booster recipients",
      colour=NULL,
      fill=NULL,
      alpha=NULL
    ) +
    theme_minimal()+
    theme(
      axis.line.x.bottom = element_blank(),
      axis.text.x.top=element_text(hjust=0),
      strip.text.y.right = element_text(angle = 0),
      axis.ticks.x=element_line(),
      legend.position = "none",
      legend.box = "horizontal"
    )+
    NULL

  ggsave(
      filename=glue("coverage_{.cohort}.png"),
      path=output_dir_qmd,
      plot=plot_coverage_cumuln,
      unit="cm",
      width=15,
      height=10
    )

  return(plot_coverage_cumuln)

}
```

#### Figure S1a: Cumulative numbers included in the matched sample for the CV cohort
```{r}
#| label: figure-coverage-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_coverage("cv")
```

#### Figure S1b: Cumulative numbers included in the matched sample for the 75+ cohort

```{r}
#| label: figure-coverage-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_coverage("age75plus")
```
&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

```{r}
#| label: table-table1-function

show_table1 <- function(.cohort){
  match_table1_wide_wide <-
    rst[[.cohort]]$match_table1_wide %>%
    filter(!(variable %in% c("N","immunossuppressed"))) %>%
    mutate(
      smd = if_else((variable %in% matching_variables$B$exact) & (matchset == "postmatch"), "-", smd)
    ) %>%
    pivot_wider(
      id_cols = c(var_label, variable, variable_levels, level),
      names_from=matchset,
      values_from = c(stat_0, stat_1, smd)
    ) %>%
    mutate(
      order_var = cumsum(lag(variable, 1, first(variable))!=variable),
      var_label = if_else(lag(as.character(var_label), n=1, default="") == as.character(var_label), "", as.character(var_label)),
    )

  match_table1_wide_wide %>%
    gt() %>%
    fmt_markdown(columns = "var_label") %>%
    tab_style(
      style = list(
        cell_fill(color = "gray96")
      ),
      locations = cells_body(
        rows = order_var%%2 == 0
      )
    ) %>%
    tab_style(
      style = list(
        cell_text(weight = "bold")
      ),
      locations = cells_column_labels(everything())
    ) %>%
    cols_label(
      var_label = "Variable",
      variable_levels = "",
      `stat_0_prematch` = glue("Pfizer BA.4-5 (N={filter(match_table1, cohort==.cohort, matchset=='prematch', variable=='N', by==0) %>% pull(n)})"),
      `stat_1_prematch` = glue("Sanofi (N={filter(match_table1, cohort==.cohort, matchset=='prematch', variable=='N', by==1) %>% pull(n)})"),
      `smd_prematch` = "SMD",
      `stat_0_postmatch` = glue("Pfizer BA.4-5 (N={filter(match_table1, cohort==.cohort, matchset=='postmatch', variable=='N', by==0) %>% pull(n)})"),
      `stat_1_postmatch` = glue("Sanofi (N={filter(match_table1, cohort==.cohort, matchset=='postmatch', variable=='N', by==1) %>% pull(n)})"),
      `smd_postmatch` = "SMD"
    ) %>%
    tab_spanner(
      label = "Before matching",
      columns = ends_with("_prematch")
    ) %>%
    tab_spanner(
      label = "After matching",
      columns = ends_with("_postmatch")
    ) %>%
    cols_align(
      align = c("right"),
      columns =  ends_with(c("_prematch", "_postmatch"))
    ) %>%
    cols_hide(c("variable", "level", "order_var")) %>%
    tab_options(
      table.font.size = 8
    )

}
```

#### Table S1a: Baseline characteristics in the CV cohort between Pfizer BA.4-5 and Sanofi groups, before and after matching

```{r}
#| label: table-table1-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_table1("cv")
```
For variables that were matched exactly the SMD is zero by design, and so are not shown.

#### Table S1b: Baseline characteristics in the 75+ cohort between Pfizer BA.4-5 and Sanofi groups, before and after matching

```{r}
#| label: table-table1-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_table1("age75plus")
```
For variables that were matched exactly the SMD is zero by design, and so are not shown.

```{r}
#| label: figure-smd-function

show_smd <- function(.cohort) {
  plot_smd <-
    rst[[.cohort]]$data_smd %>%
    filter(
      matchset %in% c("prematch", "postmatch")
    ) %>%
    filter(
      #!(variable %in% matching_variables$B$exact)
    ) %>%
    mutate(
      matchset=fct_recoderelevel(matchset, c("Pre-match"="prematch", "Post-match"="postmatch")),
    ) %>%
    droplevels() %>%
    group_by(matchset) %>%
    arrange(matchset, variable, level) %>%
    mutate(
      variable_levels = replace_na(as.character(variable_levels), ""),
    ) %>%
    mutate(
      level=fct_rev(fct_inorder(as.character(level))),
      cardn = row_number()
    ) %>%
    ggplot()+
    geom_point(aes(x=smd, y=level, group=matchset, colour=matchset), position=position_dodge(width=0.3))+
    geom_linerange(aes(xmax=smd, xmin=0, y=level, group=matchset, colour=matchset), position=position_dodge(width=0.3))+
    geom_rect(aes(alpha = variable_card, ymin = rev(cardn)-0.5, ymax =rev(cardn+0.5), group=matchset), xmin = -Inf, xmax = Inf, fill='grey', colour="transparent") +
    scale_alpha_continuous(range=c(0,0.1), guide='none')+
    scale_colour_brewer(type="qual", palette="Set1")+
    labs(
      x="Standardised mean difference\n <- more in Pfizer BA.4-5  |  more in Sanofi -> ",
      y=NULL,
      alpha=NULL,
      colour= NULL
    )+
    theme_minimal(base_size=10) +
    theme(
      legend.position = "bottom",
      strip.placement = "outside",
      strip.background = element_rect(fill="transparent", colour="transparent"),
      strip.text.y.left = element_text(angle = 0, hjust=1),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.y = element_blank(),
      panel.spacing = unit(0, "lines")
    )

  plot_smd
}

```

#### Figure S2a: Standardised Mean Differences for CV cohort between Pfizer BA.4-5 and Sanofi groups, before and after matching
```{r}
#| label: figure-smd-cv
#| out-width: '90%'
#| out-height: '100%'
#| output: asis
#| out-extra: ''

show_smd("cv")
```

#### Figure S2b: Standardised Mean Differences for 75+ cohort between Pfizer BA.4-5 and Sanofi groups, before and after matching
```{r}
#| label: figure-smd-age75plus
#| out-width: '90%'
#| out-height: '100%'
#| output: asis
#| out-extra: ''

show_smd("age75plus")
```

\newpage

```{r}
#| label: table-contrasts-function

show_contrasts <- function(.cohort){

  rst[[.cohort]]$contrasts_table %>%
    filter(
      outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    group_by(outcome_descr, subgroup_descr) %>%
    mutate(
      rd.p = if_else(row_number()==1 & (subgroup!="all"), rd.p, NA_character_),
      rr.ln.p = if_else(row_number()==1 & (subgroup!="all"), rr.ln.p, NA_character_),
      irr.ln.p = if_else(row_number()==1 & (subgroup!="all"), irr.ln.p, NA_character_)
    ) %>%
    group_by(outcome_descr) %>%
    select(
      outcome_descr,
      subgroup_descr,
      subgroup_level_descr,

      rdCI,
      rd.p,
      rrCI,
      rr.ln.p,
      irrCI,
      irr.ln.p,
    ) %>%
    mutate(
      # add indicator for row shading
      shade = cumsum((lag(as.character(subgroup_descr), 1, "")!=subgroup_descr)*1L),
      # remove repeat values of subgroup_level
      subgroup_descr = if_else(lag(as.character(subgroup_descr), n=1, default="")==subgroup_descr & row_number()!=1, "", as.character(subgroup_descr)),
    ) %>%
    ungroup() %>%
    gt(groupname_col="outcome_descr") %>%
    cols_label(

      outcome_descr = "Outcome",
      subgroup_descr = "Sub-group",
      subgroup_level_descr = "",

      rdCI = glue("Risk difference per {perN_format} people (95% CI)"),
      rd.p = "P-value",
      rrCI = "Risk ratio (95% CI)",
      rr.ln.p = "P-value",
      irrCI = "IRR (95% CI)",
      irr.ln.p = "P-value",
    ) %>%
    cols_align(
      align = c("right"),
      columns = ends_with(c("CI", ".p"))
    ) %>%
    sub_missing(
      columns = ends_with(c("CI", ".p")),
      missing_text = ""
    ) %>%
    tab_options(
      table.font.size = 8
    ) %>%
    tab_style(
      style = list(
        cell_fill(color = "gray96")
      ),
      locations = cells_body(
        rows = shade%%2 != 0
      )
    ) %>%
    cols_hide(c("shade"))

}

```

#### Table S2a: Risk differences, risk ratios, and incidence rate ratios in the CV cohort, with p-values for subgroup heterogeneity

```{r}
#| label: table-contrasts-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_contrasts("cv")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

#### Table S2b: Risk differences, risk ratios, and incidence rate ratios in the age75plus cohort, with p-values for subgroup heterogeneity

```{r}
#| label: table-contrasts-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_contrasts("age75plus")
```

\newpage

```{r}
#| label: table-fup-function

show_fup <- function(.cohort){
  rst[[.cohort]]$followup_table %>%
    filter(
      outcome %in% outcomes,
      subgroup %in% "all",
    ) %>%
    pivot_longer(
      cols=ends_with("weeks"),
      names_to = "stat",
      values_to = "value"
    ) %>%
    pivot_wider(
      id_cols = c(outcome_descr, subgroup_descr, subgroup_level_descr, treatment_descr),
      names_from = "stat",
      values_from = "value"
    ) %>%
    group_by(outcome_descr) %>%
    select(
      outcome_descr,
      #subgroup_descr,
      #subgroup_level_descr,
      treatment_descr,
      ends_with("weeks")
    ) %>%
    gt(groupname_col="outcome_descr") %>%
    cols_label(

      outcome_descr = "Outcome",
      #subgroup_descr = "Sub-group",
      #subgroup_level_descr = "",
      treatment_descr = "Vaccine",
      total_weeks = "Total",
      mean_weeks = "Mean",
      median_weeks = "Median",
      Q1_weeks = "Q1",
      Q3_weeks = "Q3",
    ) %>%
    tab_spanner(
      label = "Weeks of follow-up",
      columns = ends_with(c("weeks"))
    ) %>%
    cols_align(
      align = c("right"),
      columns = ends_with(c("weeks"))
    ) %>%
    tab_options(
      table.font.size = 8
    )
}
cat("  \n\n")
```

#### Table S3a: Summary of follow-up time for CV cohort

```{r}
#| label: table-fup-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_fup("cv")
```

#### Table S3b: Summary of follow-up time for 75+ cohort

```{r}
#| label: table-fup-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_fup("age75plus")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

```{r}
#| label: table-testrates-function

show_testingrates <- function(.cohort){

  rst[[.cohort]]$eventcounts %>%
    filter(
      #outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    transmute(
      subgroup_descr,
      subgroup_level_descr,
      treatment,
      persontime = label_number(1, 1/7)(persontime),
      test_rate = label_number(0.001, 7)(test_rate), 
    ) %>%
    pivot_wider(
      id_cols = c(subgroup_descr, subgroup_level_descr),
      names_from = treatment,
      values_from = c(persontime, test_rate)
    ) %>%
    mutate(
      # add indicator for row shading
      shade = cumsum((lag(subgroup_descr, 1, "")!=subgroup_descr)*1L),
    ) %>%
    group_by(subgroup_descr) %>%
    mutate(
      # remove repeat values of subgroup_level
      subgroup_descr = if_else(lag(subgroup_descr,n=1, default="")==subgroup_descr, "", as.character(subgroup_descr)),
      subgroup_level_descr = coalesce(subgroup_level_descr, ""),
    ) %>%
    ungroup() %>%
    gt() %>%
    cols_label(
      subgroup_descr = "Sub-group",
      subgroup_level_descr = "",
      persontime_0 = "Person-weeks",
      persontime_1 = "Person-weeks",
      test_rate_0 = "All tests",
      test_rate_1 = "All tests"
    ) %>%
    tab_spanner(
      label = "Pfizer BA.4-5",
      columns = c(persontime_0, test_rate_0)
    ) %>%
    tab_spanner(
      label = "Sanofi",
      columns = c(persontime_1, test_rate_1)
    ) %>%
    cols_align(
      align = c("right"),
      columns = c(persontime_0, persontime_1, test_rate_0, test_rate_1)
    ) %>%
    tab_style(
      style = list(
        cell_fill(color = "gray96")
      ),
      locations = cells_body(
        rows = shade%%2 != 0
      )
    ) %>%
    cols_hide(c("shade")) %>%
    tab_options(
      table.font.size = 8
    )
}

```

#### Table S4a: Testing rates after boosting for the CV cohort

```{r}
#| label: table-testingrates-cv
#| out-width: '90%'
#| out-height: '50%'
#| output: true
#| out-extra: ''

show_testingrates("cv")
```

#### Table S4b: Testing rates after boosting for the 75+ cohort

```{r}
#| label: table-testingrates-age75plus
#| out-width: '90%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

show_testingrates("age75plus")
```

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

&nbsp; &nbsp; &nbsp;

\newpage

### Cumulative incidences by subgroup

```{r}
#| label: ci-function

ci_theme <- function(...) {
  theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      #panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
      ...
    )
}

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_ci_subgroup <- function(.cohort, .subgroup){

  if(.subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_ci <- 
    rst[[.cohort]]$estimates %>%
    filter(
      subgroup==.subgroup,
      outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    mutate(
      colgroup = interaction(treatment_descr, subgroup_level_descr, sep=", "),
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
      subgroup_level_descr = fct_inorder(subgroup_level_descr)
    ) %>%
    group_by(treatment_descr, subgroup_level_descr, outcome_descr) %>%
    mutate(
      lagrisk=lag(risk,1,0),
    ) %>%
    #group_by(colgroup, outcome_descr) %>%
    ggplot(aes(group=colgroup, colour=treatment_descr, fill=treatment_descr, alpha=treatment_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    #geom_step(aes(x=lagtime, y=risk*perN, linetype=treatment_descr))+
    geom_segment(aes(x=lagtime, xend=time, y=risk*perN, yend=risk*perN, linetype=treatment_descr))+
    geom_segment(aes(x=lagtime, xend=lagtime, y=lagrisk*perN, yend=risk*perN, linetype=treatment_descr))+
    geom_rect(aes(xmin=lagtime, xmax=time, ymin=risk.ll*perN, ymax=risk.ul*perN), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), scales="free_y", switch="y")+
    scale_color_brewer(type="qual", palette="Set2", na.value="grey") +
    scale_fill_brewer(type="qual", palette="Set2", guide="none", na.value="grey") +
    #scale_color_manual(values=colourvalues, na.value="grey") +
    #scale_fill_manual(values=colourvalues, guide="none", na.value="grey") +
    scale_x_continuous(breaks = seq(0,600,28), expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_continuous(
      expand = expansion(mult=c(0,0.01)), 
      #label=label_number_risk
    )+
    scale_alpha_discrete(range=c(0.3,0.9))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Weeks",
      y=NULL,
      #title=glue("Cumulative incidence per {perN_format}"),
      colour=NULL,
      alpha=NULL,
      linetype=NULL
    )+
    guides(linetype="none", fill="none", alpha="none")+
    theme_minimal(base_size = 9)+
      theme(
        # legend.position = c(0.05,0.1),
        # legend.justification = c(0,0),
        legend.position = "bottom",
        #legend.direction = "vertical",
        axis.text.x.top=element_text(hjust=0),

        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        strip.background = element_blank(),
        strip.placement="outside",
        strip.text.y.left = element_text(angle=0),

        panel.border = element_blank(),
        panel.spacing = unit(0.8, "lines"),

      )+
      NULL

  n_subgroups <-
    rst[[.cohort]]$estimates %>%
    filter(
      subgroup==.subgroup
    ) %>% 
    `[[`("subgroup_level_descr") %>%
    n_distinct()

  # plot_combined <- patchwork::wrap_plots(
  #   plot_ci,
  #   #plot_rd,
  #   #plot_rr,
  #   #plot_irr,
  #   ncol=2,
  #   widths=c(n_subgroups,1)
  # )

  ggsave(
    filename=glue("ci_{.cohort}_{.subgroup}.png"),
    path=output_dir_qmd,
    plot=plot_ci,
  )

  plot_ci

}

```

#### Figure S3a.1: Cumulative incidence estimates per `r perN_format` people for the CV cohort

```{r}
#| label: figure-ci-cv-all
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("cv", "all")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3a.2: Cumulative incidence estimates per `r perN_format` people by age group for the CV cohort

```{r}
#| label: figure-ci-cv-ageband
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("cv", "ageband")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3a.3: Cumulative incidence estimates per `r perN_format` people by prior vaccine count for the CV cohort

```{r}
#| label: figure-ci-cv-vax_previous_group
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("cv", "vax_previous_group")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3b.1: Cumulative incidence estimates per `r perN_format` people for the 75+ cohort

```{r}
#| label:  figure-ci-age75plus-all
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("age75plus", "all")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3b.2: Cumulative incidence estimates per `r perN_format` people by age group for the 75+ cohort

```{r}
#| label: figure-ci-age75plus-ageband
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("age75plus", "ageband")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3b.2: Cumulative incidence estimates per `r perN_format` people by prior vaccine count for the 75+ cohort

```{r}
#| label: figure-ci-age75plus-vax_previous_group
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("age75plus", "vax_previous_group")
```

<!--       -->

<!--       -->

<!--       -->

#### Figure S3b.4: Cumulative incidence estimates per `r perN_format` people by clinical vulnerability for the 75+ cohort

```{r}
#| label: figure-ci-age75plus-cv
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_ci_subgroup("age75plus", "cv")
```

<!--       -->

<!--       -->

<!--       -->

```{r}
#| label: plot-contrasts-function

#colourvalues <- RColorBrewer::brewer.pal(10, "Paired")[c(1,2,5,6,3,4,7,8,9,10)]
colourvalues <- c("#D55E00", "#009E73", "#CC79A7", "#56B4E9", "#CC79A7", "#56B4E9", "#0072B2")

plot_contrast_subgroup <- function(.cohort, .subgroup, coltype){

  scale_color <- 
    if(coltype=="qual"){
      list(
        scale_color_brewer(type="qual", palette="Set1", na.value="grey"),
        scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")
      )
    } else if(coltype=="seq"){
      list(
        scale_color_viridis_d(na.value="grey"),
        scale_fill_viridis_d(guide="none", na.value="grey")
      )
    }

  if(.subgroup=="all"){
    colour_guide <- list(guides(colour="none", fill="none", alpha="none"))
  } else{
    colour_guide <- list(guides(fill="none", alpha="none"))
  }

  plot_data <- 
    rst[[.cohort]]$contrasts_daily %>%
    filter(
      subgroup==.subgroup,
      outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    mutate(
      subgroup_level_descr = fct_inorder(subgroup_level_descr)
    ) %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    mutate(
      outcome_descr = fct_relabel(outcome_descr, str_wrap, width=10),
      rr = if_else(fup_start<7, NA_real_, rr),
      rr.ll = if_else(fup_start<7, NA_real_, rr.ll),
      rr.ul = if_else(fup_start<7, NA_real_, rr.ul),
      lagrd=lag(rd,1,0),
      lagrr=lag(rr,1,1)
    )

  plot_rd <- 
    plot_data %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=0), colour="black")+
    geom_segment(aes(x=fup_start, xend=fup_end, y=rd*perN, yend=rd*perN))+
    geom_segment(aes(x=fup_start, xend=fup_start, y=lagrd*perN, yend=rd*perN))+
    geom_rect(aes(xmin=fup_start, xmax=fup_end, ymin=rd.ll*perN, ymax=rd.ul*perN), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    scale_color +
    scale_x_continuous(breaks = seq(0,600,28), minor_breaks = c(0,1,2,4)*7, expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_continuous(
      expand = expansion(mult=c(0,0.01)), 
      #=label_number(0.1, 10000)
    )+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Weeks",
      y=NULL,
      title="Cumulative\nrisk difference",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      # legend.position = c(0.05,0.1),
      # legend.justification = c(0,0),
      legend.position = "bottom",
      #legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement="outside",
      strip.text.y.left = element_text(angle=0),

      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines"),
    )+
    NULL

  plot_rr <- 
    plot_data %>%
    group_by(subgroup_level_descr, outcome_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    geom_segment(aes(x=fup_start, xend=fup_end, y=rr, yend=rr))+
    geom_segment(aes(x=fup_start, xend=fup_start, y=lagrr, yend=rr))+
    geom_rect(aes(xmin=fup_start, xmax=fup_end, ymin=rr.ll, ymax=rr.ul), alpha=0.1, colour="transparent")+
    facet_grid(rows=vars(outcome_descr), scales="free_y", switch="y")+
    scale_color +
    scale_x_continuous(breaks = seq(0,600,28), minor_breaks = c(0,1,2,4)*7, expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)))+
    coord_cartesian(xlim=c(0, NA))+
    labs(
      x="Weeks",
      y=NULL,
      title="Cumulative\nrisk ratio",
      colour=NULL
    )+
    colour_guide+
    theme_minimal(base_size = 9)+
    theme(
      legend.position="none",
      axis.text.x.top=element_text(hjust=0),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),

      #panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.text = element_blank(),
      strip.placement="outside",
      panel.border = element_blank(),
      panel.spacing = unit(0.8, "lines")
    )+
    NULL

  plot_irr <- 
    rst[[.cohort]]$contrasts_cuts  %>%
    filter(
      subgroup==.subgroup,
      outcome %in% outcomes,
      subgroup %in% subgroups,
    ) %>%
    group_by(outcome_descr, subgroup_level_descr) %>%
    ggplot(aes(group=subgroup_level_descr, colour=subgroup_level_descr, fill=subgroup_level_descr)) +
    geom_hline(aes(yintercept=1), colour="black")+
    # geom_segment(aes(x=fup_start, xend=fup_end, y=coxhr, yend=coxhr))+
    # geom_segment(aes(x=fup_start, xend=fup_start, y=lag(coxhr), yend=coxhr))+
    # geom_rect(aes(xmin=fup_start, xmax=fup_end, ymin=coxhr.ll, ymax=coxhr.ul), alpha=0.1, colour="transparent")+
    geom_point(aes(x = (fup_start+fup_end)/2, y=irr), alpha=0.5, position=position_dodge(width=5))+
    geom_linerange(aes(x = (fup_start+fup_end)/2, ymin=irr.ll, ymax=irr.ul), alpha=0.9, position=position_dodge(width=5))+
    facet_grid(rows=vars(outcome_descr), scales="free_y")+
    scale_color +
    scale_x_continuous(breaks = seq(0,600,28), minor_breaks = c(0,1,2,4)*7, expand = expansion(mult=c(0,0.01)), label = label_number(1, 1/7))+
    scale_y_log10(expand = expansion(mult=c(0,0.01)), 
                  oob=~oob_squish(.,range=log(c(1/10,10), 10)),
                  sec.axis = dup_axis(breaks=NULL, labels=NULL, name = "higher risk\n<- Pfizer BA.4-5 | Sanofi ->"))+
    coord_cartesian(xlim=c(0, 16*7))+
    colour_guide+
    labs(
      x="Weeks",
      y=NULL,
      title="Period-specific\nhazard ratio",
      colour=NULL
    )+
    theme_minimal(base_size = 9)+
    ci_theme()+
    NULL

  plot_combined <- patchwork::wrap_plots(
    plot_rd, 
    plot_rr, 
    plot_irr,
    ncol=3
  )

  ggsave(
    filename=glue("contrasts_{.cohort}_{.subgroup}.png"),
    path=output_dir_qmd,
    plot=plot_combined,
  )

  plot_combined
}

```

### Comparisons by subgroup

#### Figure S4a.1: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios for the CV cohort

```{r}
#| label: figure-contrasts-cv-all
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("cv", "all", "qual")
```

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

#### Figure S4a.2: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by age group for the CV cohort

```{r}
#| label: figure-contrasts-cv-ageband 
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("cv", "ageband", "qual")
```

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

#### Figure S4a.3: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by prior vaccine count for the CV cohort

```{r}
#| label: figure-contrasts-cv-vax_previous_group 
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("cv", "vax_previous_group", "qual")
```

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

#### Figure S4b.1: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios for the 75+ cohort

```{r}
#| label: figure-contrasts-age7plus-all
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("age75plus", "all","qual")
```

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

<!-- &nbsp; &nbsp; &nbsp; -->

#### Figure S4b.2: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by age group for the 75+ cohort

```{r}
#| label: figure-contrasts-age75plus-ageband 
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("age75plus", "ageband","qual")
```

#### Figure S4b.3: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by prior vaccine count for the 75+ cohort

```{r}
#| label: figure-contrasts-age75plus-vax_previous_group
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("age75plus", "vax_previous_group","qual")
```

#### Figure S4b.4: Cumulative risk difference, cumulative risk ratio, and period-specific incidence rate ratios by clinical vulnerability for the 75+ cohort

```{r}
#| label: figure-contrasts-age75plus-cv
#| out-width: '100%'
#| out-height: '50%'
#| output: asis
#| out-extra: ''

plot_contrast_subgroup("age75plus", "cv","qual")
```